\chapter{Conclusion}
This project has succeeded in formalising a proof that every Lawvere Theory can
be mapped to a Csystem, with the same underlying category and a length function
corresponding to the inverse of the Lawvere Theory's defining functor on
objects. As a part of this formalisation the notion of a Lawvere Theory as well
as the category of Lawvere Theories have been formalised.

The project illustrates that it is possible to carry out formalisation of
moderate scale mathematical proofs using Coq and \textit{UniMath} but also
illustrates many of the difficulties of doing so. The possibility for single
objects to have many possible names all of which are simultaneously relevant
ends up generating terms with non-trivial quantities of rewriting which can be
unintuitive to work with. Attempting to modify the formalised proof also makes
it obvious that formalisation has a high likelihood of resulting in code that is
extremely brittle, and minor modifications in one place are liable to result in
many other terms becoming ill defined. It is possible this is magnified by Coq's
use of a tactic system in addition to explicit terms to satisfy proof
obligations. The use of tactics still generates terms, which may end up being
computationally relevant, but does so opaquely and consequently can generate
terms that are difficult to reason about.

It may be interesting to compare the complexity of a similar formalisation
within alternative proof systems and libraries, such as the Coq \textit{HoTT}
library or the Agda \textit{HoTT-Agda} library, to investigate how much of this
difficulty is inherent to category theoretic formalisation in the context of
homotopy type theory, and how much is specific to either \textit{UniMath} or
Coq.