\chapter{Csystem Formalisation}
A formalisation of
Csystems\footnote{\url{https://github.com/UniMath/TypeTheory/blob/master/TypeTheory/Csystems/lCsystems.v}}
by Voevodsky already exists within the \textit{TypeTheory} library distributed
by the UniMath project. Unfortunately it defines a Csystem to be a C0-system
with additional algebraic structure, whereas we need the equivalent but simpler
definition that a Csystem is a C0-system in which the canonical squares are
pullback squares. Therefore we have separately formalised Csystems, using the
definition of a
C0-system\footnote{\url{https://github.com/UniMath/TypeTheory/blob/master/TypeTheory/Csystems/lC0systems.v}}
found in the \textit{TypeTheory} library and a newly formalised constraint that
the canonical squares be pullback squares.

We first discuss some difficulties in formalising a C0-system within Coq.
Although these have largely been dealt with by the formalisation given in
\textit{TypeTheory} their effects are still visible when working with Csystems
in Coq.

\section{Difficulties Formalising C0-systems}
Consider a canonical square of some C0-system, with $X$ and $Y$ any objects such
that $ll(X) > 0$.
\[
\begin{tikzcd}
    f^*X\ar[r,"{q(f,X)}"]\ar[d,"p_{f^*X}"]&X\ar[d,"p_X"] \\
    Y\ar[r,"f"]&ft(X)
\end{tikzcd}
\]
Within a strict dependently typed context this diagram is challenging to
directly represent, because it conceals a complication in the type of the
morphism $p_{f^*X}$. The diagram draws the morphism from $f^*X$ to $Y$, but the
type of $p_{f^*X}$ is defined to be a morphism from $f^*X$ to $ft(f^*X)$. Since
the axioms of C0-systems require that $Y=ft(f^*X)$ this is not obviously an
issue, and in paper proofs it is safe to use $Y$ and $ft(f^*X)$ interchangeably
without comment. However Coq requires us to be explicit about use of equalities
to move terms from one type to another. For instance instead of 
\lstinline|pX (f_star gt0 f)| (which represents $p_{f^*X}$) we could use
\begin{lstlisting}
transportf (fun Z => (f_star gt0 f) --> Z) (pX (f_star gt0 f)) eq
\end{lstlisting}
where \lstinline|eq| is a proof that $Y$ and $ft(f^*X)$ are equal, which changes
the type of the projection from \lstinline|f_star gt0 f --> ft (f_star gt0 f)|
to \lstinline|f_star gt0 f --> Y|. This is undesirable both because it is a
significantly larger term and because wrapping the morphism in the
\lstinline|transportf| function call makes it much harder to manipulate terms
involving it using standard lemmas about morphism composition.

An alternative approach is taken by the \textit{TypeTheory} library. For any
canonical square, a morphism is defined from $ft(f^*X)$ to $Y$, called
\lstinline|C0emor gt0 f| where \lstinline|gt0| is just the proof that $ll(X)>0$.
This morphism is exactly the identity on $Y$ transported to the desired type,
with a term similar to
\begin{lstlisting}
transportf (fun X => X --> Y) (identity Y) eq
\end{lstlisting}
where \lstinline|eq| is again an equality between $Y$ and $ft(f^*X)$.

Using this morphism, we can redraw the canonical square as follows,
\[
\begin{tikzcd}
    f^*X\ar[rr,"{q(f,X)}"]\ar[d,"p_{f^*X}"]&&X\ar[d,"p_X"] \\
    ft(f^*X)\ar[r,"C0emor"]&Y\ar[r,"f"]&ft(X)
\end{tikzcd}
\]
which is exactly equivalent except that all the morphisms have the expected
types so the paths can be written just as the composition of the stated
morphisms.

A similar issue arises when reasoning about axiom 6 of the C0-system axioms
which requires that in this canonical square,
\[
\begin{tikzcd}[sep=huge]
    {id_{ft(X)}^*X}
    \ar[r, "{q(id_{ft(X)}, X)}"]
    \ar[d, "p_{id_{ft(X)}^*X}"]&
    X\ar[d, "p_X"] \\
    ft(X)
    \ar[r, "id_{ft(X)}"]&
    ft(X)
\end{tikzcd}
\]
${id_{ft(X)}}^*X = X$ and $q(id_{ft(X)}, X) = id_X$. The second of these
constraints is hard to formalise, because $id_X$ is a morphism from $X$ to $X$,
but $q(id_{ft(X)}, X)$ comes from $id_{ft(X)}^*X$. Although these objects are
equal to even state this constraint we are required to transport the term, again
making it more difficult to work with.

\section{Pullback Csystems}
The formalisation of a pullback Csystem is as follows
\begin{lstlisting}
Definition has_canonical_pullbacks (CC : lC0system) := 
  ∏ (X Y : CC) (gt0: ll X > 0) (f : Y --> ft X), 
    isPullback 
      ((C0emor gt0 f) · f )
      (pX X) 
      (pX (f_star gt0 f))
      (q_of_f gt0 f)
      (C0ax5c gt0 f).

Definition lCsystem' := total2 (fun CC => has_canonical_pullbacks CC).
\end{lstlisting}

The name \lstinline|lCsystem'| is used to avoid collision with the definition of
\lstinline|lCsystem| in \textit{TypeTheory}. An \lstinline|lCsystem'| is a
C0-system along with a proof that its canonical squares are pullbacks.

A proof that the canonical squares are a pullback takes two objects of the
C0-system, \lstinline|X| and \lstinline|Y|, a proof that \lstinline|ll X > 0|
and a morphism from \lstinline|Y| to \lstinline|ft X| (which is enough to
specify a canonical square of the C0-system) and returns a proof that the
specified canonical square is a pullback. \lstinline|isPullback| is defined in
\textit{UniMath} and the sides of the pullback square are constructed from terms
defined in the formalisation of C0-systems. Note that where we might expect to
see \lstinline|f| in the pullback we have instead used 
\lstinline|(C0emor gt0 f)· f|.

It is also worth noting that we do not need to explicitly construct or take as
an argument a proof that the canonical squares commute, since one is
necessarily given in the C0\nobreakdash-system structure. We do not need to
worry about the choice of proof, since Csystems are small categories so the type
of morphisms between any two objects is a set and equality proofs between them
are unique.