\chapter{Background Theory}
This chapter describes the formal definitions of Lawvere Theories and CSystems,
and attempts to explain the intuition behind how these definitions model
universal algebras and dependent type theories respectively. It also briefly
describes the tools and libraries used in the formalisation work.

\section{Lawvere Theories}
This section gives the definition of a Lawvere Theory along with an intuition
for how Lawvere Theories model universal algebras and the intuitive
interpretation of the Lawvere Theory axioms. 

\subsection{Universal Algebras}
Universal algebra (CITE) provides a general description of classes of algebraic
structures with operations obeying equational laws. For example the standard
universal algebraic description of a monoid has two operations, a nullary
operation $e$ and a binary operation $\cdot$, which obey the following
equations:
\begin{gather*}
    (a\cdot b)\cdot c = a\cdot (b\cdot c) \\
    a \cdot e = a \\
    e \cdot a = a
\end{gather*}

It is important to note that no quantifiers may be used in the equations, except
implicit leading universal quantifiers over variables. Consequently the obvious
approach of using an existential quantifier to enforce the existence of an
identity element is not possible. Instead the identity element is defined as a
nullary operator rather than a member of the carrier set. However since these
ideas are equivalent we write the operator as though it were a value.

The requirement that all laws be equational unfortunately prevents the
definition of certain common algebraic structures, notably fields, within
universal algebra.

\subsection{Constructing Lawvere Theories in the Category of Sets}
To produce a Lawvere Theory corresponding to a universal algebra, we initially
consider a category with sets as objects and functions as morphisms. We will
relax this constraint later. For every natural number $n$ we define $T_n$ to be
the set of all terms constructable using the operators of the universal algebra
(in the case of monoids $e$ and $\cdot$) and $n$ distinct free variables, say
$x_0, x_1,\ldots x_{n-1}$, quotiented by the defining equations. These sets are
the objects of a category, which we call $\mathcal{T}$.

We let functions between these objects correspond to substitutions. Given a term
$t$ in $n$ variables $x_0,\ldots x_{n-1}$, we can produce a term in $m$
variables by substituting each $x_i$ in t for an element of $T_m$, so we expect
each element of $\mathcal{T}(T_n, T_m)$ to correspond to an element of
$(T_m)^n$. So we have

\[
\mathcal{T}(T_n, T_m)\cong
(T_m)^n\cong
((T_m)^1)^n)\cong
(\mathcal{T}(T_1, T_m))^n
\]
which gives us that $T_{n+m}$ must be the coproduct of $T_n$ and
$T_m$.

The category $\mathcal{T}$ has two other important properties. Firstly because
only a trivial substitution is possible on a term with no free variables it is
obvious that $T_0$ is initial. Secondly, for every function $f: \{1,\ldots n-1\}
\rightarrow \{1,\ldots m-1\}$ there is a corresponding morphism in $\mathcal{T}$
from $T_n$ to $T_m$ that for every $0\leq i\leq n-1$ substitutes $x_i$ with
$x_{f(i)}$.

\subsection{Lawvere Theories}
To capture the essential properties of $\mathcal{T}$ in a more general
categorical context, we first define the category F. Let F be the category with
objects $F_{obj}$, morphisms $F_{mor}$ and identities and composition defined as
normal for functions where
\begin{align*}
    stn(n) &= \{x\in \mathbb{N} | x < n\} \\
    F_{obj} &= \mathbb{N} \\
    F_{mor}(n, m) &= \{f | f: stn(n)\to stn(m)\}
\end{align*}

The category $F$ is isomorphic to the skeleton category of finite sets.

It is easy to show that $0$ is an initial object in $F$ and that given two
objects $n$ and $m$ the following is a coproduct structure,
\[
\begin{tikzcd}
    & n\arrow[d, "\iota_1^{n,m}"] \\
    m\arrow[r, "\iota_2^{n,m}"] & n+m
\end{tikzcd}
\]
where
\begin{align*}
    \iota_1^{n,m}:stn(n)\to& stn(n+m) \\
    x\mapsto&x \\
    \\
    \iota_2^{n,m}:stn(m)\to& stn(n+m) \\
    x\mapsto&x+n
\end{align*}
A Lawvere theory (CITE) is a pair $(T, L)$ where $T$ is a category and $L$ is a
functor $L : F \to T$ such that
\begin{enumerate}
    \item $L$ is a bijection on objects
    \item $L(0)$ is initial in $T$
    \item $L$ preserves coproducts, meaning that the diagram
    \[
    \begin{tikzcd}
        & L(n)\arrow[d, "L(\iota_1^{n,m})"] \\
        L(m)\arrow[r, "L(\iota_2^{n,m})"] & L(n+m)
    \end{tikzcd}
    \]
    is a coproduct.
\end{enumerate}

\subsection{Remark on Opposites}
The definition given for Lawvere Theories is in terms of the category $F$ and
requires that the functor $L$ preserves coproducts. In this formalisation
objects of $T$ intuitively correspond to sets of terms that can be formed with a
particular collection of variables. Under this intuition a morphism $f:A\to B$
corresponds to a function that takes a term in $A$ and produces a term in $B$ by
substituting each variable in the given term for terms in $B$.

An alternative (equivalent) formalisation uses the category $F^{op}$ and
requires that the functor $L$ preserves products. In this formalisation it
is more natural to think about objects of $T^{op}$ corresponding to contexts in
which terms can be formed (although of course unsorted universal algebra
contexts are completely specified by their length). In this case we can think
about a morphism $f:B\to A$ as a way of specifying terms for all the variables
in the context $A$ using only the variables available in context $B$ (that is,
the morphism corresponds to a way of constructing the context $A$ from the
context $B$). 

It is obvious that these two intuitions are equivalent, but some ideas are
clearer in the first intuition while others are clearer in the second one. This
distinction will become more important when we introduce Csystems, as
unfortunately much of the existing literature uses the first intuition for
Lawvere Theories and the second intuition for Csystems.

\subsection{Extracting an Algebra From a Lawvere Theory}
Given a Lawvere Theory $(T,L)$ we can produce a universal algebra that it
corresponds to. For simplicity we consider $(T^{op}, L^{op})$, and for every
natural number $n$ extract a set of $n$-ary operators $L^{op}(n,1)$. Projection
morphisms and the product structures in $T^{op}$ can be used to create morphisms
between contexts that apply operations on some members and leave others
unaffected. For instance suppose we have a morphism $f : L^{op}(2)\to L^{op}(1)$
which corresponds to an operator which we write ``$\cdot$". Writing
$\{x_0,\ldots, x_{n-1}\}$ as an intuitive representation of $L^{op}(n)$ we can
construct the following diagram in $T^{op}$
\[
\begin{tikzcd}
    &\{x_0, x_1, x_2\}
    \ar[ld, "L^{op}({\pi_{1}^{2,1}})"]
    \ar[rdd, "L^{op}({\pi_{2}^{2, 1}})"]
    & \\
    \{x_0, x_1\}\ar[dd, "f"]&& \\
    &&\{x_2\} \\
    \{(x_0\cdot x_1)\}&& \\
    &\{(x_0\cdot x_1), x_2\}
    \ar[ul, "L^{op}({\pi_{1}^{1,1}})"]
    \ar[uur, "L^{op}({\pi_{2}^{1,1}})"]
    &
\end{tikzcd}
\]
where $\pi_1^{n, m}$ and $\pi_2^{n, m}$ are the opposites of $\iota_1^{n,m}$ and
$\iota_2^{n,m}$ respectively, so by the preservation of products from $F^{op}$
to $T^{op}$ we have a unique morphism from $\{x_0, x_1, x_2\}$ to $\{(x_0\cdot
x_1), x_2\}$ that makes the diagram commute.

This technique combined with morphism composition produces operators
that correspond to the composition of other operators. An infinite set of
defining equations for the algebra are given by morphism equalities.

Notice that given the set based Lawvere Theory for monoids described earlier, an
infinite family of binary operations are generated, containing operators with
behaviour equivalent to $(x,y)\mapsto x\cdot y\cdot x$ for example. The
generated structure is equivalent to a monoid but has a very different
presentation. In fact Lawvere Theories cannot be put in bijection with
presentations of universal algebras, because up to isomorphism of categories
there is a unique Lawvere Theory for any universal algebra but many possible
presentations.

\subsection{Models of Lawvere Theories}
To give a model for the algebra represented by a Lawvere Theory, say $(T, L)$,
is to give a functor $M: T^{op}\to \catname{Set}$ which preserves terminal
objects and finite products. Consequently, on objects $M(L^{op}(1))$ picks out a
set $S$ which is the carrier set of the model. Because $M$ preserves products
$M(L^{op}(n))$ must be isomorphic to $S^n$. The operators of the model are given
by the behaviour of the functor on morphisms. For instance if the algebra
contains a nullary operator $e$, as in the case of monoids, it should correspond
to an element $e'\in T^{op}(0,1)$. Since $M(L^{op}(0))$ is a terminal set
$M(e')$ picks out a single element of $S$ as expected for a nullary operator.
Similar logic can be used to show that any $n$-ary operator in the universal
algebra is mapped by $M$ to a function $S^n\to S$.

Any such functor necessarily gives a model that respects the defining equations
of the algebra because the functor sends equal morphisms to equal
functions, and all the defining equations are represented as morphism
equalities.

\subsection{The Category Of Lawvere Theories}
We can further define relationships between different Lawvere Theories. Given
two Lawvere Theories $A = (T_1, L_1)$ and $B = (T_2, L_2)$ a morphism
$\catname{LT}(A, B)$ is a functor $G: T_1\to T_2$ such that $L_1\circ G = L_2$.

We do not explicitly require $G$ to preserve the initial object or coproducts,
but because both $L_1$ and $L_2$ are required to it must also do so.

Morphisms between Lawvere Theories correspond to translations of algebraic
structures. Similarly to how given a model of a group it is possible to
construct a model of a monoid over the same carrier set, if we have a morphism
between two Lawvere Theories $G\in \catname{LT}(A, B)$ and $M$ a model of $B$,
then we have $G^{op}\circ M$ is a model of $A$.

We can now define the category of Lawvere Theories $\catname{LT}$ with every
Lawvere Theory as an object, and the set of morphisms between two Lawvere
theories defined as above. Identities and composition in $\catname{LT}$ are
defined as in the category of small categories, and the category axioms are
therefore trivial.

\section{Csystems}
This section gives the definition of C0-systems, which are then used to give the
definition of Csystems. The intuitive correspondence between the C0-system
axioms and dependent type theory properties is briefly discussed.

\subsection{C0-systems}
A C0-system is a precursor to the definition of a Csystem, and consists of a
category, $CC$ and several pieces of additional structure:
\begin{enumerate}
    \item a function $ll: CC\to \mathbb{N}$.
    \item an object $pt\in CC$.
    \item a function $ft: CC\to CC$.
    \item for each object $X\in CC$ a morphism $p_X\in CC(X, ft(X))$.
    \item for each object $X\in CC$ and morphism $f: Y\to ft(X)$ with $ll(X) >
    0$ an object $f^*X$ and a morphism $q(f, X) : f^*X\to X$.
\end{enumerate}
which satisfy a set of axioms,
\begin{enumerate}
    \item $ll(X) = 0 \implies X=pt$.
    \item $ll(X) > 0 \implies ll(ft(X)) = ll(X) - 1$.
    \item $ft(pt) = pt$.
    \item $pt$ is a terminal object of CC.
    \item for any $X\in CC$ and $f: Y\to ft(X)$, $ft(f^*X)=Y$ and
    \[
    \begin{tikzcd}
        f^*X\ar[r,"{q(f,X)}"]\ar[d,"p_{f^*X}"]&X\ar[d,"p_X"] \\
        Y\ar[r,"f"]&ft(X)
    \end{tikzcd}
    \]
    commutes. This diagram is referred to as the canonical square of the
    C0-system.
    \item When $f$ is the identity on $ft(X)$, $f^*X = X$ and $q(f, X) = id_X$,
    so that the canonical square becomes
    \[
    \begin{tikzcd}
        X\ar[r,"id_X"]\ar[d,"p_X"]&X\ar[d,"p_X"] \\
        ft(X)\ar[r,"id_{ft(X)}"]&ft(X)
    \end{tikzcd}
    \]
    \item For $X\in CC$ with $ll(X)>0$, if we have $g:Z\to Y$ and $f:Y\to ft(X)$
    then in the following diagram created by combining two canonical squares,
    \[
    \begin{tikzcd}
        g^*(f^*X)\ar[r,"{q(g, f^*X)}"]\ar[d,"p_{g^*(f^*X)}"]&
        f^*X\ar[r,"{q(f,X)}"]\ar[d,"p_{f^*X}"]&
        X\ar[d,"p_X"] \\
        Z\ar[r,"g"]&
        Y\ar[r,"f"]&
        ft(X)
    \end{tikzcd}
    \]
    $(g\circ f)^*X = g^*(f^*X)$ and $q(g\circ f, X) = q(g, f^*X)\circ q(f, X)$.
\end{enumerate}

(CITE CSYSTEM PAPER)

The intuition for these axioms is most easily seen in terms of the contexts
intuition described previously. Each object of $CC$ corresponds to a context of
the dependent type theory being modelled, with $pt$ as the empty context,
$ll(X)$ as the number of variables in the context $X$, $ft(X)$ being the context
which is identical to $X$ but with the last variable dropped and $p_X$ is a
(trivial) mapping which forms the context $ft(X)$ given the context $X$.

Finally if we have a pair of contexts $Y$ and $X$, and a mapping which forms all
but the last element of $X$ from $Y$, encoded by the morphism $f: Y\to ft(X)$,
then we can always extend $Y$ by a single element to give a context in which is
is possible to form the context $X$. The extension of $Y$ is represented by
$f^*X$ and the morphism representing the mapping that forms $X$ is $q(f, X)$.

\subsection{Csystems}
C0-systems are not quite sufficient to model dependent type theories on their
own. When extending a context, by forming $f^*X$, there may be many possible
valid solutions, which results in multiple, non-equivalent, C0-systems for the
same dependent type theory. To prevent this we define a Csystem to be a
C0-system where $f^*X$ and $q(f, X)$ make the canonical squares into pullback
squares, intuitively requiring them to be the most general possible extension.

There is an alternative possible restriction on C0-systems to give Csystems in
terms of an additional operation and algebraic constraints upon it. The
restrictions are provably equivalent (CITE) so we deal only with the pullback
restriction here.

\section{Coq/UniMath}
\subsection{Coq}
Coq (CITE) and the UniMath library (CITE) have been used as the underlying
theory for the machine formalisation work. Coq is a well established dependently
type programming language and proof assistant based on the calculus of inductive
constructions, with many extensions to the strength of the type theory having
been added over time. Unlike Isabelle (CITE) and similar proof assistants Coq is
constructive unless non-constructive axioms are assumed. 

\subsection{\textit{UniMath}}
\textit{UniMath} is a considerably more recent project aiming to formalise the
foundations of mathematics from a Univalent (CITE) point of view. It is a large
library of Coq code containing proofs of various theorems in topics including
Category Theory, Combinatorics and KTheory. Although \textit{UniMath} is written
in Coq, it intentionally uses a restricted version of the Coq type theory, most
notably avoiding the use of general inductive types or any automated theorem
proving. Conversely, the theory is strengthened considerably by assuming the
univalence axiom (CITE), from which theorems such as function extensionality and
the negation of axiom UIP (``uniqueness of identity proofs'' which states that
for any $x$ and $x'$ if we have $y, y' : x=x'$ then $y=y'$) which are not
provable in Coq follow. This project does not make any direct use of univalence,
though function extensionality is used.

\subsection{h-levels}
Because the univalence axiom is inconsistent with the general form of axiom UIP
it is worth briefly discussing h-levels which allow us to work as though we have
a more restricted version of axiom UIP. A type $T$ is defined to be contractible
if it has exactly one inhabitant, a proof of which has type
\[iscontr(T) := \Sigma x:T.\ \Pi x':T.\ x=x'\]

A type has h-level $0$ exactly if it is contractible. A type $T$ has h-level
$n+1$ if for any two inhabitants $x, x': T$ the type of equalities between
$x=x'$ has h-level $n$.

It is fairly easy to see that types of h-level $1$ can have $0$ or $1$
inhabitants, and that the number of inhabitants of a type of h-level $2$ need
not be bounded. Types of h-level $1$ are referred to as propositions, and types
of h-level $2$ are referred to as sets. Since, by definition, equalities between
elements of a set have h-level $1$ there can be at most a single proof of the
equality, so we can work as though we have axiom UIP when working on sets.

In this project a number of lemmas regarding h-levels are used which have been
proven in \textit{UniMath}, such as: a type of h-level $n$ also has h-level
$n+1$, if types $S$ and $T$ both have h-level $n$ then $S\times T$ also has
h-level $n$ and the type of functions into a type of h-level $n$ has h-level
$n$.

\subsection{\textit{UniMath} and \textit{TypeTheory}}
The majority of the prewritten definitions and proofs used in this project comes
directly from the \textit{UniMath} library (CITE). This includes foundational
definitions like h-levels, dependent products and sums, and the natural numbers,
along with useful lemmas about these definitions and also formalisation of many
useful notions in category theory, such as categories, small categories and
functors.

However some of the output of the \textit{UniMath} project is not yet considered
sufficiently stable to be included in the main library. This content is
available in the \textit{TypeTheory} library that is also published by the
\textit{UniMath} project. A complete definition of C0-systems and Csystems is
available in this library. This project uses the C0-system definition but not
the Csystem definition because we prefer the pullback definition of a Csystem,
which is not used in the \textit{TypeTheory} library.